\blue{Numerical Investigation of Multi-Fluid Flows through Dual Saturated-Unsaturated Porous Media - change the title to something close to the main topic of the manuscript ... something with viscous flow instability or viscous fingering)}

### 2017-03-06 Papers/ViscousInstabilities/Manuscript_ViscousInstabilities_KC.tex
section 1

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\section{Introduction}\label{section:intro}
\medskip
Numerical investigation and modeling of multiphase flows in porous media have attracted the attention of the scientific community over the past 40 years. Characterization of the flow of fluids through porous media serves as the foundation of reservoir engineering studies. Oil and gas migration, transport of contaminants through groundwater are some of important physical problem that porous media flows play a crucial role. Reservoir engineers and hydro-geologists are dealing with complex geometries, irregular boundaries and heterogeneous geological formations.  Underground coal gasification, for the mining industry is an important area of interest, but more recently, also due to the global environmental problems, research focus on the CO$_{\text{2}}$ migration and trapping mechanisms in carbon capture utilization and storage (CCUS) operations~\citep{spycher_2003, chen_2006, self_2012, pruess_1990c, white_1981,jiang_2011}.

\medskip
Due to complex and/or heterogeneous geological formations, multiphase flows through porous media play an important role in understanding the afformentioned operations. Modeling of multiphase flow, is an enormous technical and computational challenge. Starting from important contributions on the field of multiphase flows from \citet{wooding_1976}, \citet{homsy_1987} and \citet{adler_1988}, in all branches of fluid mechanics, there are the instabilities, also know as flow instabilities as these, first described by \citet{saffman_1958}. The first distinction should be made between microscopic and macroscopic instabilities. Microscopic instabilities are used for phenomena occurring locally, for example, the Helmholtz and Saffman-Taylor instabilities (\citep{saffman_1959}). On the other hand the macroscopic instabilities involves two-phase flow system (\citep{bottoni_1992}). In this paper, the main focus is kept on macroscopic phenomena. A local flow instability that is the focus of this work is due to viscous and stress forces and often named as viscous instabilities or \textit{viscous fingering}.

\medskip
In large scale field operations, fingering is a relatively common condition in reservoirs with water-injection wells. As water and oil interacts, the interface between those two fluids moves along the geological formation, creating an uneven, or fingered, profile. The result of fingering is an inefficient sweeping action that can bypass significant volumes of recoverable oil and, in severe cases, an early breakthrough of water into adjacent production wellbores. This instability is mainly governed by the ratio of the mobility of the displacing fluid over that of the displaced fluid (Mobility-ratio, MR). %In a water-oil system the MR is given by eq.\ref{eq:MR}.
However, other important factors also influence the severity of viscous fingering: heterogeneous permeability fields, gravity, anisotropic dispersion, velocity dependence of dispersion, nonmonotonic viscosity profiles, etc. High mobility-ratio displacement processes introduces very complex fingering patterns.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\section{Introduction}\label{section:intro}
\medskip
Numerical investigation and modeling of multiphase flows in porous media have attracted the attention of the scientific community over the past 40 years. Characterization of the flow of fluids through porous media serves as the foundation of reservoir engineering studies. Oil and gas migration and transport of contaminants in groundwater are some of important physical problems that porous media flows play a crucial role. 
%Reservoir engineers and hydro-geologists are dealing with complex geometries, irregular boundaries and heterogeneous geological formations.
Underground coal gasification, for the mining industry is an important area of interest, but more recently, also due to the global environmental problems, most research focused on the CO$_{\text{2}}$ migration and trapping mechanisms in carbon capture utilization and storage (CCUS) operations~\citep{spycher_2003, chen_2006, self_2012, pruess_1990c, white_1981,jiang_2011}.

\medskip
Due to complex and/or heterogeneous geological formations, multiphase flows through porous media play an important role in understanding the afformentioned operations. Modeling of multiphase flow, is an enormous technical and computational challenge. Starting from important contributions on the field of multiphase flows from \citet{wooding_1976}, \citet{homsy_1987} and \citet{adler_1988}, in the research area of fluid mechanics, there are the instabilities, also know as flow instabilities as these, first described by \citet{saffman_1958}. The first distinction should be made between microscopic and macroscopic instabilities. Microscopic instabilities are used for phenomena occurring locally, for example, Kelvin-Helmholtz and Saffman-Taylor instabilities \citep{saffman_1959}. On the other hand the macroscopic instabilities involves two-phase flow system \citep{bottoni_1992}. In this paper, the test-cases taht will follow focus on macroscopic instabilities and phenomena. A local flow instability that is the focus of this work is due to viscous and stress forces and often named as viscous instabilities or \textit{viscous fingering}.

\medskip
In large scale field operations, fingering is a relatively common condition in reservoirs with water-injection wells. As water and oil interacts, the interface between those two fluids moves along the geological formation, creating an uneven, or fingered, profile. The result of fingering is an inefficient sweeping action that can bypass significant volumes of recoverable oil and, in severe cases, an early breakthrough of water into adjacent production wellbores. This instability is mainly governed by the ratio of mobility of the displacing fluid over that of the displaced fluid (Mobility-ratio, MR). %In a water-oil system the MR is given by eq.\ref{eq:MR}.
However, other important factors also influence the severity of viscous fingering: heterogeneous permeability fields, gravity, anisotropic dispersion, velocity dependence of dispersion, nonmonotonic viscosity profiles, etc. High mobility-ratio displacement processes introduces very complex fingering patterns.
